The polynomial representation of the type $A_{n - 1}$ rational Cherednik algebra in characteristic $p \mid n$

### 2015-05-19 formalwriteup.tex
Obtain dFdz2 from dFdz

Define $g(z)=\prod_{j=0}^{n-1} (1-x_jz)$ and $F(z) =\left(\sum_{k=0}^{p-1} \binom{c}{k} (g(z)-1)^k\right)$ so that $F_i(z) = \frac{F(z)}{1-x_iz}$ and $f_i = [z^p] \frac{F(z)}{1-x_iz}$.  We now show that $f_i$ are singular vectors.

\todo[inline]{Break up the proof into lemmas, e.g. for the value of $\frac{\partial F}{\partial z}$}

@@

\begin{proof}
We see easily that 
\[
\frac{\partial g}{\partial z} = g(z) \sum_j \frac{-x_j}{1-x_jz}.
\]
We now consider $\frac{\partial F}{\partial z}$. We compute
\begin{align*}
\frac{\partial F}{\partial z}%&=\frac{\partial}{\partial z}\left(\sum_{k=0}^{p-1} \binom{c}{k} (g-1)^k\right)\\
&=\left(\sum_{k=1}^{p-1}k\binom{c}{k}(g-1)^{k-1}\right)\frac{\partial g}{\partial z}\\
%&=\left(\sum_j \frac{-x_j}{1-x_jz}\right)\left(\sum_{k=1}^{p-1}k\binom{c}{k}(g-1)^{k-1}\right)g(z)\\
%&=\left(\sum_j \frac{-x_j}{1-x_jz}\right)\left(\sum_{k=0}^{p-2}(k+1)\binom{c}{k+1}(g-1)^{k}\right)g(z)\\
&=\left(\sum_j \frac{-x_j}{1-x_jz}\right)\left(\sum_{k=0}^{p-2}c\binom{c-1}{k}(g-1)^{k}\right)(g-1+1)\\
%&=\left(\sum_j \frac{-x_j}{1-x_jz}\right)\left(\left(\sum_{k=0}^{p-2}c\binom{c-1}{k}(g-1)^{k}\right)+\left(\sum_{k=0}^{p-2}c\binom{c-1}{k}(g-1)^{k+1}\right)\right)\\
&=\left(\sum_j \frac{-x_j}{1-x_jz}\right)\left(\left(\sum_{k=0}^{p-2}c\binom{c-1}{k}(g-1)^{k}\right)+\left(\sum_{k=1}^{p-1}c\binom{c-1}{k-1}(g-1)^{k}\right)\right)\\
&=\left(\sum_j \frac{-x_j}{1-x_jz}\right)\left(-c\binom{c-1}{p-1}(g-1)^{p-1}+\sum_{k=0}^{p-1}c\binom{c}{k}(g-1)^{k}\right)\\
&=\left(\sum_j \frac{-cx_j}{1-x_jz}\right)F(z)+\left(\sum_j \frac{-x_j}{1-x_jz}\right)\left(-c\binom{c-1}{p-1}(g-1)^{p-1}\right).
\end{align*}
We define 
\[
V(z)=\left(\sum_j \frac{-x_j}{1-x_jz}\right)\left(-c\binom{c-1}{p-1}(g-1)^{p-1}\right).
\]
It is then clear that 
\[
F'(z) = V(z) - \sum_{j=0}^{n-1} \frac{cx_j}{1-x_jz} F(z).
\]
Then it remains only to show that $[z^l]V(z)=0$ for $l=0,\dots,p-1$. To show this, we need only show that $V(z)=z^pv(z)$ for some power series $v(z)$.

We consider the expansion of $g(z)$; this is $g(z)=\prod_{j=0}^{n-1} (1-x_jz) = 1-z\left(\sum_i x_i\right)+z^2e_2(x_i)+\dots+z^ne_n(x_i)$ where $e_2,\dots,e_n$ are polynomials in the $x_i$. In particular we note that $\sum_i x_i=0$ by hypothesis, so $z^2 \mid g(z)-1$.

Then by the definition of $V(z)$, we see that $(g-1)^{p-1} \mid V(z)$, so $z^{2p-2} \mid V(z)$. Since $p \ge 2$, it is clear that $z^p \mid V(z)$, so we can write $V(z)=z^pv(z)$ for some power series $v(z)$ as desired. 
\end{proof}

@@

@@

\begin{lemma}\label{lem:dFdz2}
Let $y_0,\dots,y_{r-1}$ be distinct elements of $k$ and $m_i$ positive integers such that $\sum_i m_iy_i=0$.  Define 
\[
g(z)=\prod_{i=0}^{r-1} (1-y_iz)^{m_i} \text{  and } F(z)=\sum_{k=0}^{p-1} \binom{c}{k}(g-1)^k.
\]
For some $V(z)$ satisfying $[z^l]V(z)=0$ for $0 \le l \le p-1$, we have
\[
F'(z)=V(z) - \sum_j \frac{cm_jy_j}{1-y_jz}F(z).
\]
\end{lemma}
\begin{proof}
We compute as in the proof of Lemma~\ref{lem:dFdz}. We see easily that 
\[
\frac{\partial g}{\partial z} = g(z) \sum_j \frac{-m_jy_j}{1-y_jz}.
\]
We now consider $\frac{\partial F}{\partial z}$. We compute
\begin{align*}
\frac{\partial F}{\partial z}%&=\frac{\partial}{\partial z}\left(\sum_{k=0}^{p-1} \binom{c}{k} (g-1)^k\right)\\
&=\sum_{k=1}^{p-1}k\binom{c}{k}(g-1)^{k-1}\frac{\partial g}{\partial z}\\
%&=\left(\sum_j \frac{-x_j}{1-x_jz}\right)\left(\sum_{k=1}^{p-1}k\binom{c}{k}(g-1)^{k-1}\right)g(z)\\
%&=\left(\sum_j \frac{-x_j}{1-x_jz}\right)\left(\sum_{k=0}^{p-2}(k+1)\binom{c}{k+1}(g-1)^{k}\right)g(z)\\
&=- \sum_j\frac{m_jy_j}{1-y_jz}\left(\sum_{k=0}^{p-2}c\binom{c-1}{k}(g-1)^{k}\right)(g-1+1)\\
%&=\left(\sum_j \frac{-x_j}{1-x_jz}\right)\left(\left(\sum_{k=0}^{p-2}c\binom{c-1}{k}(g-1)^{k}\right)+\left(\sum_{k=0}^{p-2}c\binom{c-1}{k}(g-1)^{k+1}\right)\right)\\
&=-\sum_j\frac{m_jy_j}{1-y_jz}\left(\sum_{k=0}^{p-2}c\binom{c-1}{k}(g-1)^{k}+\sum_{k=1}^{p-1}c\binom{c-1}{k-1}(g-1)^{k}\right)\\
&=-\sum_j \frac{m_jy_j}{1-y_jz}\left(-c\binom{c-1}{p-1}(g-1)^{p-1}+\sum_{k=0}^{p-1}c\binom{c}{k}(g-1)^{k}\right)\\
&=-\sum_j \frac{cm_jy_j}{1-y_jz}F(z) + \sum_j \frac{m_jy_j}{1-y_jz}c\binom{c-1}{p-1}(g-1)^{p-1}.
\end{align*}
Defining the formal power series
\[
V(z)=\sum_j \frac{m_jy_j}{1-y_jz} c\binom{c-1}{p-1}(g-1)^{p-1},
\]
we have shown that 
\[
F'(z)=V(z)- \sum_j \frac{cm_jy_j}{1-y_jz}F(z).
\]

It remains only to show that $[z^l]V(z)=0$ for $l=0,\dots,p-1$, for which it suffices to show that $z^p \mid V(z)$.  Consider the expansion
\[
g(z)=\prod_j (1-y_jz)^{m_j} = 1-z\left(\sum_i m_iy_i \right)+z^2e_2(y_0, \ldots, y_{r - 1})+\dots+z^ne_n(y_0, \ldots, y_{r - 1})
\]
for some polynomials $e_2,\dots,e_n$. Since $\sum_i m_iy_i=0$ by hypothesis, we obtain $z^2 \mid g(z)-1$.  By definition of $V(z)$, we see that $(g-1)^{p-1} \mid V(z)$, so $z^{2p-2} \mid V(z)$. Since $p \ge 2$, this implies $z^p \mid V(z)$, as needed.
\end{proof}

\begin{lemma} \label{lem:resbyparts}
Let $y_0,\dots,y_{r-1}$ be distinct non-zero elements of $k$ and $m_i$ positive integers so that that $\sum_i m_iy_i=0$.  For $F(z)$ as in Lemma \ref{lem:dFdz2}, let $a(z)$ be the formal power series
\[
a(z)=\frac{1}{\prod_{i=0}^{r-1}(1-y_iz)}F(z).
\]
For any integer $l \le p-r$, $[z^l]a(z)$ is a linear combination of $[z^{l+1}]a(z)$, $\ldots$, $[z^{l+r}]a(z)$.
\end{lemma}
\begin{proof} 
We will use an analogue of ``residues by parts'' in a formal power series setting.   We begin by showing that 
\[
a'(z)=z^pb(z)-\sum_j \frac{(m_jc-1)y_j}{1-y_jz}a(z)
\]
for some formal power series $b(z)$. By Lemma~\ref{lem:dFdz2} we have
\[
F'(z)=V(z)+\left(\sum_j \frac{-cm_jy_j}{1-y_jz}F(z)\right),
\]
where $[z^l]V(z)=0$ for $0 \le l \le p-1$. 
We now compute
\begin{align*}
a'(z)&=\frac{d}{dz}\left(\frac{1}{\prod_{i=0}^{r-1}(1-y_iz)}F(z)\right)\\
&=\frac{1}{\prod_{i=0}^{r-1}(1-y_iz)}F'(z)+\sum_{j=0}^{r-1}\frac{y_j}{(1-y_jz)\prod_{i=0}^{r-1}(1-y_iz)}F(z)\\
&=\frac{1}{\prod_{i=0}^{r-1}(1-y_iz)}\left(V(z)+\sum_j \frac{-cm_jy_j}{1-y_jz}F(z)\right)+\sum_{j=0}^{r-1}\frac{y_j}{1-y_jz}a(z)\\
&=\frac{V(z)}{\prod_{i=0}^{r-1}(1-y_iz)}+\sum_j \frac{-cm_jy_j}{1-y_jz}a(z)+\sum_{j=0}^{r-1}\frac{y_j}{1-y_jz}a(z)\\
&=\frac{V(z)}{\prod_{i=0}^{r-1}(1-y_iz)}-\sum_j \frac{(m_jc-1)y_j}{1-y_jz}a(z).
\end{align*}
Since $[z^l]V(z)=0$ for $0 \le l \le p-1$ we see that 
\[
\frac{V(z)}{\prod_{i=0}^{r-1}(1-y_iz)}=z^pb(z)
\]
for some power series $b(z)$, giving the claim.

>>>>

\subsection{Definition of the singular vectors}

Define the polynomials 
\[
g(z)=\prod_{j=0}^{n-1} (1-x_jz) \qquad \text{ and } \qquad F(z) =\left(\sum_{k=0}^{p-1} \binom{c}{k} (g(z)-1)^k\right).
\]
In terms of $g(z)$ and $F(z)$, we have $F_i(z) = \frac{F(z)}{1-x_iz}$ and $f_i = [z^p] \frac{F(z)}{1-x_iz}$.  We will show that $f_i$ are singular vectors.

\subsection{Computation of some partial derivatives}

@@

\begin{proof}
We see easily that 
\[
\frac{\partial g}{\partial z} = g(z) \sum_j \frac{-x_j}{1-x_jz}.
\]
We now consider $\frac{\partial F}{\partial z}$. We compute
\begin{align*}
\frac{\partial F}{\partial z}%&=\frac{\partial}{\partial z}\left(\sum_{k=0}^{p-1} \binom{c}{k} (g-1)^k\right)\\
&=\sum_{k=1}^{p-1}k\binom{c}{k}(g-1)^{k-1}\frac{\partial g}{\partial z}\\
%&=\left(\sum_j \frac{-x_j}{1-x_jz}\right)\left(\sum_{k=1}^{p-1}k\binom{c}{k}(g-1)^{k-1}\right)g(z)\\
%&=\left(\sum_j \frac{-x_j}{1-x_jz}\right)\left(\sum_{k=0}^{p-2}(k+1)\binom{c}{k+1}(g-1)^{k}\right)g(z)\\
&=-\sum_j \frac{x_j}{1-x_jz}\sum_{k=0}^{p-2}c\binom{c-1}{k}(g-1)^{k}(g-1+1)\\
%&=\left(\sum_j \frac{-x_j}{1-x_jz}\right)\left(\left(\sum_{k=0}^{p-2}c\binom{c-1}{k}(g-1)^{k}\right)+\left(\sum_{k=0}^{p-2}c\binom{c-1}{k}(g-1)^{k+1}\right)\right)\\
&=-\sum_j \frac{x_j}{1-x_jz}\left(\sum_{k=0}^{p-2}c\binom{c-1}{k}(g-1)^{k}+\sum_{k=1}^{p-1}c\binom{c-1}{k-1}(g-1)^{k}\right)\\
&=-\sum_j \frac{x_j}{1-x_jz}\left(-c\binom{c-1}{p-1}(g-1)^{p-1}+\sum_{k=0}^{p-1}c\binom{c}{k}(g-1)^{k}\right)\\
&=-\sum_j \frac{cx_j}{1-x_jz}F(z)+\sum_j \frac{x_j}{1-x_jz}c\binom{c-1}{p-1}(g-1)^{p-1}.
\end{align*}
Defining the formal power series
\[
V(z)=\sum_j \frac{x_j}{1-x_jz}c\binom{c-1}{p-1}(g-1)^{p-1},
\]
we see that 
\[
F'(z) = V(z) - \sum_{j=0}^{n-1} \frac{cx_j}{1-x_jz} F(z).
\]
It remains only to show that $[z^l]V(z)=0$ for $l=0,\dots,p-1$.  Consider the expansion 
\[
g(z)=\prod_{j=0}^{n-1} (1-x_jz) = 1-z\left(\sum_i x_i\right)+z^2e_2(x_0, \ldots, x_{n - 1})+\dots+z^ne_n(x_0, \ldots, x_{n - 1})
\]
for some polynomials $e_2, \ldots, e_n$.  Recalling that $\sum_i x_i=0$ by hypothesis, we see that $z^2 \mid g(z)-1$. By definition of $V(z)$, we see that $(g-1)^{p-1} \mid V(z)$, hence $z^{2p-2} \mid V(z)$. Since $p \ge 2$, this implies $z^p \mid V(z)$, completing the proof.
\end{proof}

@@

\subsection{Proving $f_i$ are singular vectors}

@@

\begin{corollary}\label{corr:dFdz2}
Let $y_0,\dots,y_{r-1}$ be distinct elements of $k$ and $m_i$ positive integers such that $\sum_i m_iy_i=0$.  Define 
\[
g(z)=\prod_{i=0}^{r-1} (1-y_iz)^{m_i} \text{  and } F(z)=\sum_{k=0}^{p-1} \binom{c}{k}(g-1)^k.
\]
For some $V(z)$ satisfying $[z^l]V(z)=0$ for $0 \le l \le p-1$, we have
\[
F'(z)=V(z) - \sum_j \frac{cm_jy_j}{1-y_jz}F(z).
\]
\end{corollary}
\begin{proof}
Evaluate Lemma \ref{lem:dFdz} by substituting $m_i$ copies of $y_i$ for the $x_i$.
\end{proof}

\begin{lemma} \label{lem:resbyparts}
Let $y_0,\dots,y_{r-1}$ be distinct non-zero elements of $k$ and $m_i$ positive integers so that that $\sum_i m_iy_i=0$.  For $F(z)$ as in Corollary \ref{corr:dFdz2}, let $a(z)$ be the formal power series
\[
a(z)=\frac{1}{\prod_{i=0}^{r-1}(1-y_iz)}F(z).
\]
For any integer $l \le p-r$, $[z^l]a(z)$ is a linear combination of $[z^{l+1}]a(z)$, $\ldots$, $[z^{l+r}]a(z)$.
\end{lemma}
\begin{proof} 
We will use an analogue of ``residues by parts'' in a formal power series setting.   We begin by showing that 
\[
a'(z)=z^pb(z)-\sum_j \frac{(m_jc-1)y_j}{1-y_jz}a(z)
\]
for some formal power series $b(z)$. By Corollary~\ref{corr:dFdz2} we have
\[
F'(z)=V(z)+\left(\sum_j \frac{-cm_jy_j}{1-y_jz}F(z)\right),
\]
where $[z^l]V(z)=0$ for $0 \le l \le p-1$. 
We now compute
\begin{align*}
a'(z)&=\frac{d}{dz}\left(\frac{1}{\prod_{i=0}^{r-1}(1-y_iz)}F(z)\right)\\
&=\frac{1}{\prod_{i=0}^{r-1}(1-y_iz)}F'(z)+\sum_{j=0}^{r-1}\frac{y_j}{(1-y_jz)\prod_{i=0}^{r-1}(1-y_iz)}F(z)\\
&=\frac{1}{\prod_{i=0}^{r-1}(1-y_iz)}\left(V(z)+\sum_j \frac{-cm_jy_j}{1-y_jz}F(z)\right)+\sum_{j=0}^{r-1}\frac{y_j}{1-y_jz}a(z)\\
&=\frac{V(z)}{\prod_{i=0}^{r-1}(1-y_iz)}+\sum_j \frac{-cm_jy_j}{1-y_jz}a(z)+\sum_{j=0}^{r-1}\frac{y_j}{1-y_jz}a(z)\\
&=\frac{V(z)}{\prod_{i=0}^{r-1}(1-y_iz)}-\sum_j \frac{(m_jc-1)y_j}{1-y_jz}a(z).
\end{align*}
Since $[z^l]V(z)=0$ for $0 \le l \le p-1$ we see that 
\[
\frac{V(z)}{\prod_{i=0}^{r-1}(1-y_iz)}=z^pb(z)
\]
for some power series $b(z)$, giving the claim.
